---
title: "AREC 736 D  HW!"
author: "Innocent Vomitadyo"
date: "2022-10-17"
output: word_document
---
``



#1. Simulate a dataset with one X variable and one Y variable in R or Stata that does not contain a confounder or a collider. Let the relationship between Y and X be 5, that is the population parameter ğ›½5 = 5 (you can set ğ›½0 to anything you want). Assuming X causes Y, recover the coefficient on X using OLS.

```{r setup, include=FALSE}
rm(list =ls())
X <- floor(runif(1000, min=0, max=45))
E <- rnorm(1000, mean=0, sd=0.01)
B0 <- 2
B1 <- 5
Y <- B0 + B1*X + E
df <- data.frame(Y, X)
head(df)
summary(df)

```



#Recover the coefficient on X using OLS.

```{r}
OLS1 <- lm(Y~X, data=df)
OLS1
summary(OLS1)
```
#########################################################################################################

#2. Alter your program from step 1 to include a confounder (a variable Z that is correlated with the X variable and the error term in your model)




```{r}
Z <- floor(runif(1000, min=0, max=45))
X <- 4+ 0.6*Z + rnorm(1000, mean=0, sd=0.1)
B0 <- 2
B1 <- 5
B2 <- 3
Y <- B0 + B1*X + B2*Z + rnorm(1000, mean=0, sd=0.1)
df1 <- data.frame(Y, X, Z)
head(df1)
summary(df1)

```

#########################################################################################################
#2a Estimate a model of Y on X using OLS. Report your coefficient. 


```{r}
OLS2 <- lm(Y~X, data=df1)
OLS2
summary(OLS2)

```

#Coefficients have changes. B0 is now ------- and B1 is now ------

#########################################################################################################




#2a(i) What should it have been?

#estimates should be the same as in 1. We want B0 to be equal to 2 and B1 = 5


#########################################################################################################

#2a (ii) What is the bias?

#Bias B1 = 5-  9.996728 = 
#Bias B0 - 2- -17.954363 = 

#Bias due to Endogeneity 
#The missing variable is now appearing in the error term and  the X variable is correlated with the error term


#########################################################################################################

#2(b) Estimate a model of Y on X but include your confounder. Is the bias eliminated? 

```{r}

OLS3 <- lm(Y~X + Z, data=df1)
OLS3
summary(OLS3)

```
#The Bias is eliminated.
#The estimates are close to 5 but are not exactly equal.  

#Why is the coefficient not exactly equal to 5?
#Becaue of the error term.  

#########################################################################################################

# 2c. Provide an example of an omitted variable that could cause bias in a regression.

Omitted variable is comes when we have a variable that has a significant impact on Y and correlated with X  is left out of the equation. For example,  A variable A causes Y and A also depends on X.  







#########################################################################################################

#2(d). Provide an example of selection bias that could cause bias in a regression. Is this the same as omitted variable bias?

Selection bias happens when the sample is not random such that the subset is not representative of the target population. 
Examples include  
 -when participants from a specified population are more likely to be selected to participate in a trial than others
 -when the method used to assign subjects to the study groups is inadequate and produces systematic differences between the        participants in the study groups.
 when we have Loss to follow-up and attrition. 

This not the same as omitted variable bias. 

##################.#######################################################################################


#2(e). Provide an example of non-random measurement error in an X variable that could lead to bias. Is this the same as omitted variable bias?








#########################################################################################################
